Scalar Items in Embedded Position: {A}n Experimental Revisit

### 2012-08-07 text/paper/paper1.tex
discussion sec com

High scores of success in our general control condition clearly
indicate that participants understood the incremental verification
task and were in general able to give judgements at the earliest
possible point in a picture sequence. Results obtained for the
preference-related controls moreover show that our design is able to
detect multiple readings and preferences among these. Crucially, the
amount of response types for \lc- and \ec-readings reflected the known
preference for the former, \emph{irrespective} of the order in which
each reading could first be judged in a given sequence. Given this,
the best and most straightforward explanation of the observed data is
the hypothesis that the actual preferences among readings for both
\as- and \es-sentences alike are:
\begin{exe}
  \ex \label{bsp:observed-preferences} \lit \ > \ \loc \ \ ( > \glb)
\end{exe}
where it is unclear whether the global readings are entirely
unavailable or simply severely dispreferred. Crucially, our data does
attest local readings, albeit that they are not preferred across the
board but preferred, as it seems, only by a small, but substantial
number of consistent localists.

This is bad news for both traditionalism and grammaticalism, as no
version of either would predict the pattern in
(\ref{bsp:observed-preferences}). The main problem is that
traditionalism, if left on its own, fails to predict local readings,
while grammaticalism, when it relies on the strongest meaning
hypothesis, fails to predict the preference order altogether. What are
the problems exactly? Could either position be aided by innocuous
auxiliary assumptions to be compatible with
(\ref{bsp:observed-preferences})?

It is not a major obstacle for traditionalism to predict that global
readings are available. The weak variety of traditionalism holds that
the extra assumptions needed to derive global readings are not
standardly available, and so traditionalism is compatible with a lack
of observed global readings. Alternatively, traditionalists could
argue that the task's rather high processing demands lead to a
suppression of global implicature readings, in line with the findings
of \citet{NeysDe-NeysSchaeken2007:When-People-Are} that working memory
load negatively affects the number of scalar implicature responses in
sentence verification tasks. Unfortunately, to the extent that global
readings are explained away as dispreferred or unavailable,
traditionalism also looses the possibility of accounting for local
readings of \as-sentences in its unrestricted variety.\fn{Remember
  that what we called unrestricted traditionalism can account for
  local readings of \as-sentences by assuming that a sentence of the
  form \emph{Some of the $X$ are related to some of the $Y$} is a
  feasible alternative.} This is because unrestricted traditionalism
explains local and global readings of \as-sentences in essentially the
same manner. On top of that, traditionalism cannot explain local
readings for \es-sentences at all. In sum, traditionalism alone cannot
explain (\ref{bsp:observed-preferences}), but needs an additional
explanation of local readings for \emph{both} \as- \emph{and}
\es-sentences.

Grammaticalism, on the other hand, fails to predict the preference
relation in (\ref{bsp:observed-preferences}), because these strictly
contradict the strongest meaning hypothesis in either of the two
varieties suggested by \citet{ChierchiaFox2008:The-Grammatical}. This
shows most clearly in the fact that both versions of the strongest
meaning hypothesis predict that local readings of \as-sentences are
preferred over literal ones. The problem, therefore, seems to be the
selection of readings by the strongest meaning hypothesis. In fact, if
(\ref{bsp:observed-preferences}) is the actual preference over
readings, then the strongest meaning hypothesis is also conceptually
on the wrong track, because no account that merely looks at the
logical strengths of readings can plausibly predict the preference
pattern in (\ref{bsp:observed-preferences}) for both \as- and
\es-sentences at the same time. (Try, if you please!) Turning away
from comparisons of logical strength, a selection criterion that
produces (\ref{bsp:observed-preferences}) is ready at hand: if we
assume that each insertion of $\exh(\cdot)$ operator is \emph{costly}
anti-proportional to the depth of its embedding, then we derive a
preference of literal readings over local ones over global ones. But
whether such an economy principle is sound and in line with further
empirical data is an interesting matter of further investigation. What
it does not explain in any case, is why some subjects in our study
gave local responses \emph{consistently}, while others did not. In
sum, in as far as grammaticalism depends on the strongest meaning
hypothesis, its predictive success is equally thin as that of
traditionalism.

\bigskip

On the face of it, there is a very clear winner among the stylized
candidate theories that we introduced in
Section~\ref{sec:theories-predictions}. Assuming that global readings
are not available at all, the preference pattern in
(\ref{bsp:observed-preferences}) is exactly what \emph{literalism}
would predict. Remember that literalism was the conventionalist
position which posited a lexical ambiguity of scalar \emph{some} with
a preference for the semantically weaker meaning \emph{some and maybe
  all} over the upper-bounded reading \emph{some but not all}.

@@

>>>>

High scores of success in our general control condition clearly
indicate that participants understood the incremental verification
task and were in general able to give judgements at the earliest
possible point in a picture sequence. Results obtained for the
preference-related controls moreover show that our design is able to
detect multiple readings and preferences among these. Crucially, the
amount of response types for \lc- and \ec-readings reflected the known
preference for the former, \emph{irrespective} of the order in which
each reading could first be judged in a given sequence. Given this,
the best and most straightforward explanation of the observed data is
the hypothesis that the actual preferences among readings for both
\as- and \es-sentences alike are:
\begin{exe}
  \ex \label{bsp:observed-preferences} \lit \ > \ \loc \ \ ( > \glb)
\end{exe}
where it is unclear whether the global readings are entirely
unavailable or simply severely dispreferred. Crucially, our data
attests local readings independently of a stress accent on the
embedded scalar item, albeit that local readings are not preferred to
literal readings and given, as it seems, only by a small, but
substantial number of consistent localists.

This is bad news for both traditionalism and grammaticalism, as no
version of either would predict the pattern in
(\ref{bsp:observed-preferences}). The main problem is that
traditionalism, if left on its own, fails to predict local readings,
while grammaticalism, when it relies on the strongest meaning
hypothesis, fails to predict the right preference order. Let's look
more closely at each position in turn.

As for traditionalism, it is not a major obstacle for this position to
predict that global readings are available. The weak variety of
traditionalism holds that the extra assumptions needed to derive
global readings are not standardly available, and so traditionalism is
compatible with a lack of observed global readings. Alternatively,
traditionalists could argue that the task's rather high processing
demands lead to a suppression of global readings, in line with the
findings of \citet{NeysDe-NeysSchaeken2007:When-People-Are} that
working memory load negatively affects the number of scalar
implicature responses in sentence verification tasks. Unfortunately,
to the extent that global readings are explained away as dispreferred
or unavailable, traditionalism also looses the possibility of
accounting for local readings of \as-sentences in its unrestricted
variety.\fn{Remember that what we called unrestricted traditionalism
  can account for local readings of \as-sentences by assuming that a
  sentence of the form \emph{Some of the $X$ are related to some of
    the $Y$} is a feasible alternative.} This is because unrestricted
traditionalism explains local and global readings of \as-sentences in
essentially the same manner. On top of that, traditionalism cannot
explain local readings for \es-sentences. Since our design presented
sentence material auditorily, the traditionalist cannot resort to
effects of silent intonation to explain these local readings
away. Local readings showed on the neutral accentuation case as
well. Notice that it is not a problem for this argument that the local
responses were not emphasized with a contrastive stress on the
embedded scalar. In sum, traditionalism alone cannot explain
(\ref{bsp:observed-preferences}), but needs an additional explanation
of local readings for \emph{both} \as- \emph{and} \es-sentences.

Grammaticalism, on the other hand, fails to predict the preference
relation in (\ref{bsp:observed-preferences}), because these strictly
contradict the strongest meaning hypothesis in either of the two
varieties suggested by \citet{ChierchiaFox2008:The-Grammatical}. This
shows most clearly in the fact that both versions of the strongest
meaning hypothesis predict that local readings of \as-sentences are
preferred over literal ones. The problem, therefore, seems to be the
selection of readings by the strongest meaning hypothesis. In fact, if
(\ref{bsp:observed-preferences}) is the actual preference over
readings, then the strongest meaning hypothesis is also conceptually
on the wrong track, because no account that merely looks at the
logical strengths of readings can plausibly predict the preference
pattern in (\ref{bsp:observed-preferences}) for both \as- and
\es-sentences at the same time. (Try, if you please!)

It may seem as if grammaticalism is better suited to cope with the
data than traditionalism, because traditionalism cannot cope with
local readings for principled reasons while the only thing that
grammaticalism got wrong is the selection criterion over preferred
readings, i.e., it mistakenly put all its coin on the strongest
meaning hypothesis. Turning away from comparisons of logical strength,
a \emph{prima facie} plausible selection criterion that produces
(\ref{bsp:observed-preferences}) is ready at hand: if we assume that
each insertion of $\exh(\cdot)$ operator is \emph{costly}
anti-proportional to the depth of its embedding under other logical
operators, then we derive a preference of literal readings over local
ones over global ones. Whether such an \emph{economy principle} is
sound and in line with further empirical data would then be an
interesting matter of further investigation. On the entirely
speculative supposition that it did turn out to be a good predictor,
grammaticalism would be saved in a straightforward way. But is this an
argument for grammaticalism contra traditionalism? No it clearly is
not. The crucial point is that grammaticalism needs a selection
criterion, for without one it is not a predictive theory at all, and
that the only one that has been seriously proposed fails to predict
the observed data. That grammaticalism can do with many such selection
criteria is not, as such, an argument in its favor over
traditionalism. The case remains that both theories fail on the
observed data.

There is a variation of an existing theory of scalar implicature that
does explain the preference ordering (\ref{bsp:observed-preferences})
in a completely straightforward manner. The theory of interest is
\emph{conventionalism} (also sometimes called \emph{lexicalism})
\citep[e.g.][]{LevinsonPresumptiveMeanings2000,Chierchia:2004_ScalarImplicatures}.
The general idea is that if the upper-bounding inference from
\emph{some} to \emph{some but not all} occurs frequently, then it
would be inefficient to execute the Gricean reasoning that
traditionalists propose over and over again. Conventionalism therefore
assumes that scalar inferences have become routinized, hard-wired, so
much so that there are two lexical entries for a scalar item like
\emph{some}, one with logical meaning \emph{some and maybe all} and
one with the upper-bounded meaning \emph{some but not
  all}. \citet{LevinsonPresumptiveMeanings2000} proposed that the
latter is the preferred default meaning and has therefore ofttimes be
addressed as \mymark{defaultism}. Unfortunately, defaultism has strong
empirical evidence against it, because several processing studies show
that the computation of scalar inferences appears to be time- and
effort-consuming, when they \emph{do} arise, not when they
\emph{don't}, as defaultism would have it
\citep[c.f.][]{BrehenyKatsos2006:Are-Generalised,BrehenyKatsos2008:Experimental-In,NeysDe-NeysSchaeken2007:When-People-Are}. But
there is also competing evidence that scalar inferences are available
in immediate, low-level processes, such as the resolution of
referential expressions in a visual-world paradigm
\citep{GrodnerKlein2010:Some-and-Possib}. For our purposes, it is
interesting to note that the preference relation in
(\ref{bsp:observed-preferences}) is predicted by the dual of
defaultism, call it \emph{literalism}, which posits a lexical
ambiguity but predicts that the logical meanings are \emph{ceteris
  paribus} preferred over the upper-bounded meanings.

Unfortunately, there are problems with conventionalism that seem to
require that it be backed up by traditionalism
\citep[e.g.][]{Sauerland2012:The-Computation}.\fn{\citet{Sauerland2012:The-Computation}
  eventually dismisses even conventionalism backed up by
  traditionalism in favor of grammaticalism based on cases of
  so-called \emph{intermediate implicatures}, where what seems to be
  necessary is the insertion of an exhaustivity operator in
  intermediate scope position \citep[see][for
  details]{Sauerland2012:The-Computation}. Appreciation of such
  intermediate implicatures requires more argumentation, involving a
  discussion of \emph{Hurford's constraint}
  \citep[c.f.][]{FoxSpector:Economy-and-Emb,ChierchiaFox2009:Hurfords-Constr,ChierchiaFox2008:The-Grammatical}.}
For instance, conventionalism cannot account for \emph{indirect
  implicatures}, such as the inference from
(\ref{bsp:indirect-SI-Target}) to (\ref{bsp:indirect-SI-Inference})
which cannot be explained by positing a lexical ambiguity for
\emph{all}.

\begin{exe}
  \ex \label{bsp:indirect-SI}
    \begin{xlist}
      \ex \label{bsp:indirect-SI-Target} Hans did not solve all of the problems.
      \ex \label{bsp:indirect-SI-Inference} Hans solved some of the problems.
    \end{xlist}
\end{exe}

\noindent In the context of the present study, this would then be good
news for traditionalism, because a combined
traditionalism+conventionalism would predict the observed data
well. The combined position would even be able to explain something
that grammaticalism cannot make sense of as easily, namely the
observation that there were rather consistent localists among the
participants of our study. The combined tradionalism+conventionalism
could explain this by the assumption that some (lazy?, smart?)
speakers have made a conventionalized upper-bounding meaning entry for
scalar \emph{some} while others keep computing things on the
fly. It is not clear how grammaticalism can account for the
intra-subjective consistency.

In sum, if the interpretation of our data as suggested here is correct
(we will argue presently that it is), then both traditionalism and
grammaticalism are in need of revision. Traditionalism must explain
the availability of local readings, grammaticalism must revise its
selection criterion among readings.

@@

\begin{itemize}
\item future research: isolate consistent localists to probe into \emph{processing}
\end{itemize}
